### 2020-07-10 Thesis-Chapter2-Background.tex
Add 4thAm in Flux notes
Address a number of 'fixme's; improve writing

Cryptography is the technical foundation for many forms of computer and network security, at every scale. Security is
often defined in terms of several properties: authentication, integrity, non-repudiation, confidentiality, availability,
and authorization \cite{shostack_threat_2014}. Security properties are systems properties, but systems properties emerge
from the composition of their parts, and innumerable primitives and technologies rely on cryptography.
% FIXME: reading note: paragraph makes sense, but need to read it multiple times for it to sink in. Improve.

% Spoofing                  Authentication
% Tampering                 Integrity
% Repudiation               Non-Repudiation
% Information Disclosure    Confidentiality
% Denial of Service         Availability
% Elevation of Privilege    Authorization

Cryptographic primitives include symmetric key encryption, symmetric key encryption, one-way hash functions, and digital
signatures. A few technologies that directly rely on these primitives include \ac{TLS}, \ac{PKI}, \ac{SSO} (e.g.
Kerberos), \acp{VPN}, secure password storage, disk encryption tools, and secure- and trusted-boot processes. A few
systems that directly rely on these technologies include the local and global financial systems; industrial control
systems and critical infrastructure; health information systems; private, government, and military communications; and
the integrity of the internet as a whole.
% FIXME: Jumping rather quickly from primitives to huge systems. Perhaps scale it more smoothly. Connect little
% primitives to higher technologies, higher technologies to underpinnings of network computer systems, and network
% computer systems as critical to these industries.

% Let us break security down into three scales: primitives, technologies, and systems. A primitive is a conceptual
% building block that serves a single technical function. A technology is a set of implemented primitives glued together
% to provide useful features. A system is a collection of technologies that collectively serve the ultimate outcome.
% The security properties critical at the systems level depend

% Primitives: one-way hashes enable integrity checks, block ciphers enable efficient confidential data storage, and
% digital signature enable non-repudiation

% Technologies: Transport Layer Security (TLS) enables confidentiality and integrity over the internet, Kerberos enables
% convenient authentication, and virtual private networks (VPNs) enable confidential and authenticated remote network
% connections

If the cryptographic primitives that compose these systems were to be compromised, the systems they comprise could
collapse. If the financial system loses integrity, health information systems lose confidentiality, critical
infrastructure loses availability, government communications lose authenticity and non-repudiation, or the internet as a
whole loses integrity, inconceivably large damage would occur.

% FIXME: Kristen doesn't like the wording
Encryption's dual role as enabler of strong privacy and cornerstone of computer security may be the central conflict in
the EA debate. Those who focus on its role in privacy object to its absolute power. Those who focus on its role in
security rely on its absolute power.

@@

Computerized encryption is a new technology, but the field of cryptography is old, since demand for privacy is as old as
communication itself. The most well known ancient example of rudimentary encryption is the Caesar cipher, named after
the character transposition technique Julius Caesar used to protect private correspondence
\cite{luciano_cryptology_1987}. Significant developments include the first formal cryptographic study by Arab scholars
in the eighth century, advancements made out of necessity in the twentieth century's world wars, and the application of
computers to cryptographic problems \cite{kahn_codebreakers_1996}. Claude Shannon's formalized the modern ``mathematical
analysis of cryptography'' in 1949 \cite{shannon_communication_1949}, and Whitfield Diffie and Martin Hellman published
research on \ac{public-key-crypto} in 1976 \cite{diffie_new_1976}.

The discovery of \ac{public-key-crypto} was an important advance. Combined with computer networking and personal
computing advances in the 1980s, it for the first time put the power of strong encryption not just in the hands of
governments and militaries, but of ordinary people. However, this power enables very strong privacy, and privacy is a
property governments aim to regulate. The use of encryption for for private purposes has a contentious history. In 1587,
Mary Queen of Scots was convicted of treason based on evidence from decrypted letters, and in 1807, prosecutors trying
Aaron Burr for treason tried to force testimony from his secretary on the contents of encrypted messages
\cite{kerr_encryption_2017}. Government's natural stance towards privacy makes it inherently uncomfortable with the
rapid increase in availability of strong encryption.
% FIXME: The point is made---these developments made gov'ts uncomfortable (just look at these examples!)---but the line
% of thought is not clear enough to the reader.

@@

Despite the failure of the \ac{clipperchip}, the debate over access to strong encryption continued, primarily focused on
various key escrow proposals \cite{thompson_2015}.
% FIXME: export controls were a hot topic of the time, too. Can't just drop that thread.
In 1996, two pro-encryption bills were introduced in the US Congress.
In the Senate, it was S.1726, the \ac{PROCODEAct} of 1996, to abolish the export controls on encryption software
\cite{burns_s1726_1996}. In the House, it was HR.3011, the \ac{SAFEAct} of 1996, to explicitly allow arbitrarily strong
encryption for all legal activity \cite{goodlatte_hr3011_1996}. Indeed, the \ac{SAFEAct} went further, including
``provisions that would have barred the government from creating a mandatory key escrow system and would also have
removed export restrictions on most generally available software and hardware containing encryption''
\cite{kehl_right_2015}. Zimmermann, the author of PGP, who had by this time endured an investigation by the Customs
Service for publishing his work, testified before the Senate in favor of the \ac{PROCODEAct} \cite{zimmermann_1996}. In
November of that year, the Clinton administration released an executive order effectively removing export controls on
strong encryption products along the lines of the \ac{PROCODEAct}, which never came to pass \cite{clinton_1996}.

@@

The second development was the events of and reaction to the morning of September 11, 2001. When the U.S. had already
been expanding police surveillance activities to combat communism, crime, drugs, and terrorism, the scale and audacity
of the September 11 attacks seemed to justify yet greater investigatory powers \cite{bloss_escalating_2007}. October of
2001 saw passage of the USA PATRIOT Act, which had several significant impacts: it weakened the limitation for using
\ac{FISA} requests on U.S. citizens; it expanded the scope of what could be compelled via \ac{FISA} orders; it
% FIXME: introduce FISA above, in the "had already been expanding" section. Then we have more of a narrative of
% expanding power.
authorized ``roving'' wiretaps; and it increased the power of \acp{NSL}, which can be used without judicial review to
compel information from digital service providers while precluding any public disclosure of the event
\cite{sensenbrenner_2001} \cite{shamsi_2011}. The government was performing more surveillance than ever, and even these
weakened limitations would be repeatedly violated \cite{shamsi_2011} \cite{tucker_2020}.

@@

\paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as
antagonistic, an assumption implicit in in the ``nothing-to-hide'' argument, but discussions in recent years have shown
% FIXME: expand on "nothing to hide"? that the debate is more about ``Security vs. Security'' than ``Security vs.
Privacy'' \cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security: public security, the
pursuit of law enforcement, and cybersecurity, the pursuit of the technical community. Both sides of the debate have by
now acknowledged that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public
security vs. cybersecurity---which in return affects public security---may be more important \cite{schneier_2019}.

@@

The following sections summarize the encryption regulatory environment and technical approaches to \ac{EA}.
% FIXME: remove this sentence?

@@

% IDEA: Create an appendix to go into greater detail?
% Could go into these, as well as the commented-out ones below
%  Wiretap Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_119.html)
%  Stored Communications Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_121.html)
%  Pen Register Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_II_20_206.html)

>>>>

Cryptography is the technical foundation for many forms of computer and network security. Security is defined in terms
of several properties: authentication, integrity, non-repudiation, confidentiality, availability, and authorization
\cite{shostack_threat_2014}. These system-level properties emerge from both the composition of the system's components
(architecture) and the security of the components themselves; likewise, the security of each component emerges from the
architecture of its subcomponents and the security of the subcompenents themselves. At the bottom of this analysis, the
security of primitive components commonly relies on cryptography.

\mytab{table:security-and-crypto} connects each security property to its cryptographic basis. While lacking direct
cryptographic foundations, availability indirectly relies on each other property, and authorization schemes are
typically rooted in authentication.

% TODO: clarify relationship between encryption and cryptography?

% Spoofing                  Authentication
% Tampering                 Integrity
% Repudiation               Non-Repudiation
% Information Disclosure    Confidentiality
% Denial of Service         Availability
% Elevation of Privilege    Authorization
\begin{table}[h]
    \caption[Security Properties and Cryptography]{The Cryptographic Basis of Security Properties}
    \label{table:security-and-crypto}
    \centering
    \begin{tabular}{|l|l|}
        \hline
        \thead{Property} & \thead{Cryptographic Basis} \\ \hline
        Authentication  & Digital certificates \\ \hline
        Integrity       & Cryptographic hashes \\ \hline
        Non-repudiation & Digital signatures \\ \hline
        Confidentiality & Encryption \\ \hline
        Availability    & \small{\ii{based on architecture}} \\ \hline
        Authorization   & \small{\ii{rooted in authentication}} \\ \hline
    \end{tabular}
\end{table}

% Cryptographic primitives include symmetric key encryption, symmetric key encryption, one-way hash functions, and digital
% signatures. A few technologies that directly rely on these primitives include \ac{TLS}, \ac{PKI}, \ac{SSO} (e.g.
% Kerberos), \acp{VPN}, secure password storage, disk encryption tools, and secure- and trusted-boot processes. A few
% systems that directly rely on these technologies include the local and global financial systems; industrial control
% systems and critical infrastructure; health information systems; private, government, and military communications; and
% the integrity of the internet as a whole.
% % If I bring this back---this is jumping rather quickly from primitives to huge systems. Scale it more smoothly.
% % Connect little primitives to higher technologies, higher technologies to underpinnings of network computer systems,
% % and network computer systems as critical to these industries.

Violations of cryptographic integrity in primitive components, introduced by design or by accident, could have
catastrophic effects. Compromised digital certificates, hashes, or encryption would enable spoofing, tampering, and
information disclosure done both for their own sake and as tools in general attacks. The importance of taking these
tools away is clear when such attacks include interference with elections \cite{mueller_2018}, multi-billion dollar
disruptions of business \cite{greenberg_2018_notpetya}, and hospital network ransoms that disrupt care of patients
\cite{goodin_ransomware_2019}.

% NOTE: tried to find examples of high profile bugs or hacks that relied on crypto failures, but most are still
% architectural (exposed databases, default credentials (Mirai botnet)), human (BEC), or insecure code (buffer
% overflows--key to EternalBlue (wannacry and notpetya) and MS08-067 (Conficker)).

Encryption's dual role as enabler of privacy and cornerstone of security is at the heart of the EA debate. Cryptography
has an absolute, mathematical power---a power that is objectionable in the realm of privacy, but necessary in the realm
of security.

@@

Computerized encryption is a new technology, but the field of cryptography is old, since demand for privacy is as old as
communication itself. The most well known ancient example of rudimentary encryption is the Caesar cipher, named after
the character transposition technique Julius Caesar used to protect private correspondence
\cite{luciano_cryptology_1987}. Significant developments include the first formal cryptographic study by Arab scholars
in the eighth century, advancements made out of necessity in the twentieth century's world wars, and the application of
computers to cryptographic problems \cite{kahn_codebreakers_1996}. Claude Shannon's formalized the modern ``mathematical
analysis of cryptography'' in 1949 \cite{shannon_communication_1949}, and Whitfield Diffie and Martin Hellman published
research on \ac{public-key-crypto} in 1976 \cite{diffie_new_1976}. The discovery of \ac{public-key-crypto} was an
important advance. Combined with computer networking and personal computing advances in the 1980s, it for the first time
put the power of strong encryption not just in the hands of governments and militaries, but of ordinary people.

However, this power enables very strong privacy, and privacy is a property governments aim to regulate. The use of
encryption for for private purposes has a contentious history. In 1587, Mary Queen of Scots was convicted of treason
based on evidence from decrypted letters, and in 1807, prosecutors trying Aaron Burr for treason tried to force
testimony from his secretary on the contents of encrypted messages \cite{kerr_encryption_2017}. Government's natural
stance towards privacy makes it inherently uncomfortable with the rapid increase in availability of strong encryption.

@@

Despite the failure of the \ac{clipperchip}, the debate over export controls and access to strong encryption continued,
primarily focused on various key escrow proposals \cite{thompson_2015}. In 1996, two pro-encryption bills were
introduced in the US Congress. In the Senate, it was S.1726, the \ac{PROCODEAct} of 1996, to abolish the export controls
on encryption software \cite{burns_s1726_1996}. In the House, it was HR.3011, the \ac{SAFEAct} of 1996, to explicitly
allow arbitrarily strong encryption for all legal activity \cite{goodlatte_hr3011_1996}. Indeed, the \ac{SAFEAct} went
further, including ``provisions that would have barred the government from creating a mandatory key escrow system and
would also have removed export restrictions on most generally available software and hardware containing encryption''
\cite{kehl_right_2015}. Zimmermann, the author of PGP, who had by this time endured an investigation by the Customs
Service for publishing his work, testified before the Senate in favor of the \ac{PROCODEAct} \cite{zimmermann_1996}. In
November of that year, the Clinton administration released an executive order effectively removing export controls on
strong encryption products along the lines of the \ac{PROCODEAct}, which never came to pass \cite{clinton_1996}.

@@

The second development was the events of and reaction to the morning of September 11, 2001. These events occurred in the
context of decades of all three branches of government expanding law enforcement powers. Presidents promoted the idea of
being ``tough on crime'' and launched a war on drugs; legislators passed bills such as 1978's \ac{FISA}, which provided
broad surveillance powers over foreign nationals, and in some cases U.S. citizens \cite{rodino_1978}; and the judiciary
practiced a ``jurisprudence of crime control'' that gave police broad leeway \cite{gizzi_fourth_2016}.

In this context of expanded surveillance, the scale and audacity of the September 11 attacks seemed to justify yet
greater investigatory powers \cite{bloss_escalating_2007}. October of 2001 saw passage of the USA PATRIOT Act, which had
several significant impacts: it weakened the limitation for using \ac{FISA} requests on U.S. citizens; it expanded the
scope of what could be compelled via \ac{FISA} orders; it authorized ``roving'' wiretaps; and it increased the power of
\acp{NSL}, which can be used without judicial review to compel information from digital service providers while
precluding any public disclosure of the event \cite{sensenbrenner_2001} \cite{shamsi_2011}. The government was
performing more surveillance than ever, and even these weakened limitations would be repeatedly violated
\cite{shamsi_2011} \cite{tucker_2020}.

@@

\paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as
antagonistic, an assumption implicit in in the ``nothing-to-hide'' argument, but discussions in recent years have shown
that the debate is more about ``Security vs. Security'' than ``Security vs. Privacy''
\cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security: public security, the pursuit of
law enforcement, and cybersecurity, the pursuit of the technical community. Both sides of the debate have by now
acknowledged that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public
security vs. cybersecurity---which in return affects public security---may be more important \cite{schneier_2019}.

@@

@@

% IDEA: Create an appendix to go into greater detail?
% Could go into these, as well as the commented-out ones below
%  Wiretap Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_119.html)
%  Stored Communications Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_121.html)
%  Pen Register Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_II_20_206.html)
% TODO: I _really_ need to understand these and how they're used today.
